Make any stupid plot you want

### 2021-01-29 draft/sections/implementation.tex
start of reworked implementation

\documentclass[../main.tex]{subfiles}
\begin{document}

\section{Matplottoy}

We build on the existing Matplotlib architecture \cite{hunterMatplotlib2DGraphics2007} so that we can initially focus on the data to graphic transformations and rely on Matplotlib for the other graphical elements of the visualization and the rendering. We first introduce an implementation of scatter, line, and bar charts because they map to the fundemental marks of point, line, and area. We then introduce aggregated bar charts to show how to build more complex graphics.

\subsection{Scatter}
 A scatter plot \cite{friendlyBriefHistoryData2006a,tukeyExploratoryDataAnalysis1977} is a chart type for plotting discrete values against each other. Minimally a scatter plot requires and x or y position, but the other variables in our dataset can be mapped to visual aspects of the scatter graphical mark. 
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
    \label{fig:scatter}
\end{figure}

For the scatter plot in figure~\ref{fig:scatter}, we define $Q$ as 
 \begin{equation}
    Q(\mu_{xpos}, \mu_{ypos}, \mu_{facecolor}, \mu_{s})
 \end{equation}

which we implement building on the \texttt{matplotlib.collections} API:

\begin{minted}{python}
    class Point(mcollections.Collection):
        # this is the visual fiber 
        required = {'x', 'y'}
        optional = {'facecolors', 's'} 
        def __init__(self, data, transforms, *args, **kwargs):
            """
            Parameters
            ----------
            data: sections of the fiber bundle
            transforms
            """
            super().__init__(*args, **kwargs)

            # check that the data you're trying to transform 
            # has a way to provide vertex data  
            assert 'vertex' in data.FB.K['tables']
            # check that you've given the required parameters
            utils.check_constraints(Point, transforms.keys())
            utils.validate_transforms(data.FB.F, transforms)

            self.data = data
            self.transforms = transforms

        def draw(self, renderer, *args, **kwargs):
            view = self.data.view('vertex') #resolve to size

            # call tau step
            visual = utils.convert_transforms(view, self.transforms)

            # assembles taus to generate idiom
            # dictionary here in place of visual(parameter) function

            visual['s'] = itertools.repeat(visual.get('s', 0.05))
            visual['facecolors'] = visual.get('facecolors', "C0")
            #switch out to a marker 
            self._paths = [mpath.Path.circle(center=(x,y), radius=s)  
                            for (x, y, s) 
                            in zip(visual['x'],visual['y'], visual['s'])]

            self.set_facecolors(visual['facecolors'])
            super().draw(renderer, *args, **kwargs)
    \end{minted}

>>>>

\documentclass[../main.tex]{subfiles}

\begin{document}
%what, why, how
\section{Prototype Implementation: Matplottoy}
\label{sec:implementation}
\begin{figure}[H]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/line_1.png}
    \end{subfigure}
    \caption{Scatter plot and line plot implemented using prototype artists and data models, building on Matplotlib rendering.}
    \label{fig:code_scatter_line}
\end{figure}

We implemented a proof of concept of the topological model to demonstrate its feasibility. We implemented the artist classes for the scatter, line, and bar plots shown in figure~\ref{fig:code_scatter_line} because scatter and line have different continuities \gbase\ and visual bundles \vtotal\. We built our prototype on top of the existing Matplotlib architecture \cite{hunterMatplotlib2DGraphics2007,hunterArchitectureOpenSource} so that we can initially focus on the data to graphic transformations and rely on Matplotlib for the rendering.

To generate figure~\ref{fig:code_scatter_line} in our prototype, we set up a canvas \mintinline{python}{fig} and an axes to draw on \mintinline{python}{ax} and 
\begin{multicols*}{2}
\begin{minted}[linenos]{python}
    fig, ax = plt.subplots()
    artist = Point(table, transforms)
    ax.add_artist(artist)
\end{minted}
\columnbreak
\begin{minted}[linenos]{python}
    fig, ax = plt.subplots()
    artist = Line(table, transforms)
    ax.add_artist(artist)
\end{minted}
\end{multicols*}
 and then pass in \dsection=\mintinline{python}{table} and \vchannel=\mintinline{python}{transforms} into equivalence classes \vartisteq=\mintinline{python}{Point} and  \vartisteq=\mintinline{python}{Line}.

\subsubsection{Artist Class $\vartist^{\prime}$}
What, why, how>
We implement the \vartist\ functions that transform data to visual representations as the equivalence class of artists such that the user configure the visual sections \vsection\ as a parameter \mintinline{python}{transforms}

\begin{minted}[linenos]{python}
    class ArtistClass(matplotlib.artist.Artist):
        def __init__(self, data, transforms, *args, **kwargs):
            super().__init__(*args, **kwargs)
        def draw(self, renderer, *args, **kwargs):
            super().draw(renderer, *args, **kwargs)
\end{minted}

As mentioned in section~\ref{sec:artist_equivalance}, it would be impractical to implement graph types as \vartist\ functions where all the visual parameters are specified.

\subsubsection{Encoders \vchannel}
what, why, how?
\subsubsection{Data \dtotal}
what, why, how?

\subsubsection{Case Study: Iris}

### 2021-01-29 draft/sections/math.tex
start of reworked implementation

\subsubsection{Equivalance class of artists $\vartist^{\prime}$}
As formulated above, every artist function \vartist\ has fixed \vchannel\ and \vmark\ which  generates a distinct graphic \gsection. It is impractical to implement an artist for every single graphic; instead we implement the equivalence class of artists $\{\vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}\}$. Equivalent artists have the same fiber bundle \vtotal\ and same assembly function \vmark\, but act on different sections \vsection. To further simplify implementation, we identify a minimal \vfiber\ associated with each $\vartist^{\prime}$ that defines what visual characteristics of the graphic must originate in the data\note{needs citation, maybe friendlys history or acquired codes of meaning}. 
\begin{figure}
    \caption{Each of these graphics is generated by a different artist \vartist which is the equivalance class of scatter plots $\vartist^{\prime}$\note{this is gonna be a whole bunch of scatter plots}}
    \label{fig:artist_equivalence}
\end{figure}
For example, a scatter plot of red circles is the output of one artist, a scatter plot of green squares the output of another, as are the rest of the graphs in figure~\ref{fig:artist_equivalance}. These two artists are equivalent since their only difference is in the literal visual encodings (color, shape). Shape and color could also be defined in \vmark\, but the position must come from the fiber $\vfiber=(xpos,ypos)$ since fundementally a scatter plot is the plotting of one position against another\cite{friendlyBriefHistoryData2008}. We also use this criteria to identify derivative types, for example the bubble chart\cite{tufteVisualDisplayQuantitative2001} is a type of scatter where by definition the glyph size is mapped from the data.

>>>>

\subsubsection{Equivalance class of artists $\vartist^{\prime}$}
\label{sec:artist_equivalance}
As formulated above, every artist function \vartist\ has fixed \vchannel\ and \vmark\ which  generates a distinct graphic \gsection. It is impractical to implement an artist for every single graphic; instead we implement the equivalence class of artists $\{\vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}\}$. Equivalent artists have the same fiber bundle \vtotal\ and same assembly function \vmark\, but act on different sections \vsection. To further simplify implementation, we identify a minimal \vfiber\ associated with each $\vartist^{\prime}$ that defines what visual characteristics of the graphic must originate in the data\note{needs citation, maybe friendlys history or acquired codes of meaning}. 
\begin{figure}
    \caption{Each of these graphics is generated by a different artist \vartist which is the equivalance class of scatter plots $\vartist^{\prime}$\note{this is gonna be a whole bunch of scatter plots}}
    \label{fig:artist_equivalence}
\end{figure}
For example, a scatter plot of red circles is the output of one artist, a scatter plot of green squares the output of another, as are the rest of the graphs in figure~\ref{fig:artist_equivalance}. These two artists are equivalent since their only difference is in the literal visual encodings (color, shape). Shape and color could also be defined in \vmark\, but the position must come from the fiber $\vfiber=(xpos,ypos)$ since fundementally a scatter plot is the plotting of one position against another\cite{friendlyBriefHistoryData2008}. We also use this criteria to identify derivative types, for example the bubble chart\cite{tufteVisualDisplayQuantitative2001} is a type of scatter where by definition the glyph size is mapped from the data.
